What's in a name? On the (automated) topic naming of software maintenance activities

### 2010-03-02 icpc/icpc2010.tex
methodology change

\subsection{Topic and Concept Extraction}
Topic analysis uses tools such as Latent Dirichlet Allocation (LDA)~\cite{Blei2003} and Latent Semantic Indexing (LSI) to extract
independent word distributions (topics) from
	documents (commit log comments). Many researchers~\cite{marcus04wcre,Poshyvanyk2007,lukins08wcre,Linstead2007} have applied tools like LSI and LDA for mining repositories, in particular analyzing source code, bugs or developer comments.
\begin{comments}
Ideally these extracted topics 
correlate with actual development topics that the developers discussed
during the development of the software project. For example, our technique might identify a collection of words: ``list next has iterator add'', which we then identify (manually) as concerning the topics of collections and the Iterator pattern.
\end{comments}

Typically a topic analysis tool like LDA will try to find $N$ indepdent word distributions found from within the word distributions of all the messages.

In topic analysis a single document, such as a commit message, can be related to multiple topics. Representing documents as a mixture of topics maps well to commits to source code repositories, which often have more than one purpose.  A topic
represents both a word distribution and a group of commit log comments
that are related to each other by their content.  In this paper a topic
is a set of tokens extracted from commit messages found within a
project's source control system (SCS).

In this paper we use LDA to identify topics.  LDA works by creating topic models for each period. A topic model is a statistical summary of the important subjects for that dataset.  The input to the algorithm is a set of documents, with contents treated as a bag of words, and a parameter to set the number of topics to generate. The algorithm returns a word-document matrix relating the occurrence frequency of a 
 The idea is similar to other IR approaches like TF/IDF (term frequency/inverse document frequency), which find uncommon words that aren't found in other wordlists. These approaches are unsupervised in that they rely solely on the source data with no human intervention.

@@

\subsection{Supervised learning}
While unsupervised techniques (LSI and LDA are both unsupervised) are appealing in their lack of human intervention, supervised learners have the advantage of domain knowledge. This typically results in improved results. In supervised learning, the data is divided into slices. One slice is manually annotated by the domain expert

@@

We wanted to compare two projects with open, accessible source code in similar functional areas. This allows us to partially control for differences in functional requirements. We selected MySQL and MaxDB, open-source, near-commercial quality database software. MaxDB started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project has 940 thousand lines of C source code\footnote{generated using David A. Wheeler's `SLOCCount'.}. There are approximately ... developers working on it. MySQL 3.23, released in early 2001, contains 320 thousand lines of C and C++ source code. The MySQL project started in 1994.

>>>>

\subsection{Topic and Concept Extraction}
Topic extraction, sometimes called concept extraction, uses tools such as Latent Dirichlet Allocation (LDA)~\cite{Blei2003} and Latent Semantic Indexing (LSI) to extract
independent word distributions (topics) from
	documents (commit log comments). Many researchers~\cite{marcus04wcre,Poshyvanyk2007,lukins08wcre,Linstead2007} have applied tools like LSI and LDA for mining repositories, in particular analyzing source code, bugs or developer comments.
\begin{comments}
Ideally these extracted topics 
correlate with actual development topics that the developers discussed
during the development of the software project. For example, our technique might identify a collection of words: ``list next has iterator add'', which we then identify (manually) as concerning the topics of collections and the Iterator pattern.
\end{comments}

Typically a topic analysis tool like LDA will try to find $N$ independent word distributions found within the word distributions of all the messages. Linear combinations of these $N$ word distributions are then meant to be able to recreate the word distributions of all of the underlying messages. These $N$ word distributions effectively form topics: cross cutting collections of words relevant to 1 or more documents. Our problem is that these topics are not easy to interpret and we feel that automatic labelling or naming of these topics would be helpful with respect to interpreting what a topic is about. LDA's topics extracted in an unsupervised manner, LDA relies solely on the source data with no human intervention.

In topic analysis a single document, such as a commit message, can be related to multiple topics. Representing documents as a mixture of topics maps well to commits to source code repositories, which often have more than one purpose~\cite{hindle09icsm}.  A topic represents both a word distribution and a group of commit log comments that are related to each other by their content.  In this paper a topic is a set of tokens extracted from commit messages found within a project's source control system (SCS).

\begin{comment}
In this paper we use LDA to identify topics.  LDA works by creating topic models for each period. A topic model is a statistical summary of the important subjects for that dataset.  The input to the algorithm is a set of documents, with contents treated as a bag of words, and a parameter to set the number of topics to generate. The algorithm returns a word-document matrix relating the occurrence frequency of a 
 The idea is similar to other IR approaches like TF/IDF (term frequency/inverse document frequency), which find uncommon words that aren't found in other wordlists. These approaches are unsupervised in that they rely solely on the source data with no human intervention.
\end{comment}

@@

\subsection{Supervised learning}
While unsupervised techniques (LSI and LDA are both unsupervised) are appealing in their lack of human intervention, supervised learners have the advantage of domain knowledge. This typically results in improved results. In supervised learning, the data is divided into slices. One slice is manually annotated by the domain expert
In this paper, we employ the WEKA and Mulan machine learning frameworks in order to test supervised learning.

@@

% XXX how many do we have
We wanted to compare two mature software projects, that had open repositories, within the same domain. This allows us to partially control for differences in functional requirements. We selected MySQL and MaxDB, open-source, near-commercial quality database software. MaxDB started in the late 1970s as a research project, and was later acquired by SAP. As of version 7.500, released April 2007, the project has 940 thousand lines of C source code\footnote{generated using David A. Wheeler's `SLOCCount'.}. There are approximately ... developers working on it. MySQL 3.23, released in early 2001, contains 320 thousand lines of C and C++ source code. The MySQL project started in 1994.
